# Colophon

This document was produced using `R`, R Markdown, \LaTeX. The document was built using the `bookdown` package for `R`. The document template is a variation on TJ Mahr's [`buckydown`](https://github.com/tjmahr/buckydown) template, which was itself an adaptation of documents designed by and for students at the Universities of Washington and Wisconsin. The PDF is typeset using pdf\TeX. The body text is $\text{\emph{\familydefault}}$ in 12pt size.

The data and source code for this dissertation have been organized into an online [Git repository on Bitbucket](https://bitbucket.org/mikedecrescenzo/dissertation). A hard copy of the thesis can be found in the University of Wisconsin library system.

- Git repository: <https://bitbucket.org/mikedecrescenzo/dissertation>
- `huskydown`: <https://github.com/benmarwick/huskydown>
- `bookdown`: <https://github.com/rstudio/bookdown>
<!-- [uwthesis1]: http://staff.washington.edu/fox/tex/
[uwthesis2]: https://github.com/UWIT-IAM/UWThesis
[ucbthesis]: https://github.com/stevenpollack/ucbthesis
[dissertate]: https://github.com/suchow/Dissertate -->

This version of the thesis was generated on `r Sys.time()`. The repository is
currently at this commit:

```{r echo = FALSE, include = TRUE}
# I've set eval=FALSE to ensure Travis-CI can run
# if you're not using Travis-CI, then eval=TRUE will be fine
library(git2r)
if (in_repository()) {
  summary(commits()[[1]]) 
} else { 
  message("We are not in a git repository") 
}
```

<!-- The computational environment that was used to generate this version is as
follows:

```{r echo = FALSE}
devtools::session_info()
```
 -->

