# Modeling the Constituency's Policy Preferences {#ch:model}

<!-- bold math symbol -->
```{block, include = knitr::is_html_output(), cache = FALSE}
$\newcommand{\bm}[1]{\boldsymbol{\mathbf{#1}}}$
```



```{r knitr-02-model, include = FALSE, cache = FALSE}
source(here::here("assets-bookdown", "knitr-helpers.R"))
```

```{r r-02-model, cache = FALSE}
library("knitr")
library("here")
library("magrittr")
library("tidyverse")
library("extrafont")
library("latex2exp")
```



To study how partisan constituencies are represented in primary elections, we require a measure of the partisan constituency's policy preferences. This chapter presents the statistical model that I use to estimate the policy ideal points of district-party publics. 

This chapter proceeds in three major steps. First, I review the theoretical basis for ideal point models, which can be traced to spatial models of policy choice from classic formal theory work in American political science such as @downs:1957:economic-theory. I connect these formal models to statistical models of policy ideal points [in a style that follows @clinton-jackman-rivers:2004:ideal] as well as their connection to Item Response Theory (IRT) models from psychometrics and education testing [e.g. @fox:2010:bayesian-IRT].

Second, I specify and test the group-level model that I build and employ in my analysis of district-party publics. This discussion includes details that are relevant to Bayesian estimation, including identification restrictions on the latent policy space, specification of prior distributions, and model parameterizations that expedite estimation with Markov chain Monte Carlo (MCMC). I begin with a static model for one time period, and then I describe a dynamic model that smooths estimates across time using hierarchical priors for model parameters [@caughey-warshaw:2015:DGIRT]. I test both the static and the dynamic models by fitting them to simulated data and determining how well they recover known parameter values.
<!------- TO DO ---------
- Prior checking
------------------------->

Lastly, I describe how I fit the model to real data. This section describes data collection, data processing, and model performance, and it includes a preview of the estimates before a descriptive analysis in Chapter&nbsp;\@ref(ch:descriptives).


<!------- TO DO ---------
- Should this include a comparison to CDW measures? Hill measures?
------------------------->



## Spatial Models and Ideal Points

<!-- goal: spatial >> IRT >> Parameterizations -->

Ideal points are constructs from *spatial* models of political choice. These models exist under formal theory---they simplify scenarios in the political world into sets of actors whose behaviors obey utility functions that conform to mathematical assumptions. Spatial models invoke a concept of "policy space," where possible policy outcomes are represented as locations along a number line or a multidimensional plane. A canonical example is a left-right continuum, where progressive or "liberal" policies occupy locations on the left side of the continuum, while conservative policies are on the right side. Depending on the scenario, actors of various types (voters, legislators, executives...) have to choose policy outcomes that result from their interactions with other players. Actors are at least partially motivated by their policy preferences, so they strive to achieve policy outcomes that they prefer. Sometimes these actors' face constrained choices---they can't achieve their most-preferred policy, so they settle on something that is as close as they can get.

```{r spatial-data}
dd <- tibble(
  Left = -1,
  Right = 1,
  Actor = 0.3
) 

long_dd <- dd %>%
  gather(key = label, value = location) %>%
  mutate(height = 0) %>%  
  print()

x_min <- -2
x_max <- 2
u_scale <- 2

utility <- tibble(
  x = seq(x_min, x_max, .01),
  utility_loss = u_scale * -(x - dd$Actor)^2,
  Right = u_scale * -(x - dd$Right)^2,
  Left = u_scale * -(x - dd$Left)^2
) %>%
  print(n = nrow(.))

```


```{r plot-space-1, out.width = "100%", fig.height = 2, fig.width = 8, include = TRUE, fig.cap = "An Actor and two policy outcomes (Left and Right) represented as locations in ideological/policy space"}
utility_plot <- ggplot(long_dd) +
  aes(x = location, y = height) +
  geom_point() +
  theme_void() +
  annotate(geom = "line", x = c(-2, 2), y = 0) +
  geom_text(aes(label = label), nudge_y = 1) +
  coord_cartesian(ylim = c(-3, 3)) 

utility_plot
```

Figure&nbsp;\@ref(fig:plot-space-1) plots a simple example of an Actor's choice over two policies in one-dimensional policy space. The "Left" outcome is a more progressive policy outcome than the "Right" outcome, indicated by their locations on the line. The Actor has a location herself, which corresponds to her most-desired policy outcome. There is no policy located exactly at the Actor's preferred location, but the Actor is closer to the Right policy than to the Left. Supposing that the Actor could make an error-free choice over which policy to implement, it appears she would prefer the Right outcome to the Left outcome.

Formal models are more careful to specify the assumptions governing these scenarios, which can be complicated in many cases. For example, we can imagine that locations along the left-right continuum can be assigned values on the real number line. Figure&nbsp;\@ref(fig:plot-space-1) shows a one-dimensional number line, but we could generalize the scenario to say that policy locations can be represented by locations $\mathbb{R}^{d}$, or a real-valued coordinate in $d$-dimensional space. The Actor's location is synonymous with her "ideal point," her most-preferred policy. This is the point where the Actor's utility, in an economic utility model, is maximized with respect to policy considerations. Utility implies that the Actor has a utility function that is defined over the policy space; the Actor's realized utility depends on the distance between her ideal point and a potential policy outcome. Outcomes nearer to the Actor's ideal point are generally more preferred than farther outcomes, but this too is subject to assumptions about the shape of the Actor's utility function. Typically utility functions are assumed to be symmetric around an Actor's ideal point, so the "closer" policy is always more preferred, all else equal. The notion of an ideal point is similar to a "bliss point" in microeconomics: the quantity of a good consumed such that *more* consumption would result in *less* overall utility. Whether an Actor can choose the closest policy to herself depends on the structure of the game: the presence and strategy profiles of other Actors, the sequence of play, and the presence of other non-policy features of Actors' utility functions.

<!-- to do: $d$ cite -->
<!-- canonical cite for the or the notion of policy dimensions? -->

Formal models of ideal points are distinct from statistical models of ideal points. Formal models are primarily theoretical exercises; they explore the incentives and likely actions of Actors in specific choice contexts, building theoretical intuitions that can be applied in the study of real-world politics with real data. Statistical models, on the other hand, explicitly or implicitly *assume* a formal model as given and estimate its parameters using data. Data could come from legislators casting voting on bills, judges ruling on case outcomes, survey respondents stating their policy preferences (as in this project), and other situations. Researchers are typically interested in parameter estimates for the Actors' ideal points, although sometimes the parameters about the policy alternatives are substantively interesting.
<!------- TO DO ---------
- when are item params interesting  
------------------------->

<!------- TO DO ---------
- describe where ideal point ≠ ideology?
- ideal point is an operationalization of a nebulous, potentially high-dimensional concept called ideology?
- latent construct: none of the parameters are known, we have identifiability issues, and the model is still assumed!!
- this really is model-driven, we are assuming a model
------------------------->

Having distinguished formal and statistical models, I now show a derivation of a statistical model from a formal model. This exercise model will serve as a theoretical basis for the class of statistical models explored in this dissertation. I begin with notation to describe an arbitrary number of actors indexed $i \in \{1, \ldots, n\}$ making an arbitrary number of policy choices (bills, survey items, etc.) indexed $j \in \{1, \ldots, J\}$. Every Actor has an ideal point, or a location in the policy space, represented by $\theta_{i}$. Every task is choice between a Left policy located at $L_{j}$ and a Right policy located at $R_{j}$. 

The utility that an Actor receives from a Left or Right choice is a function of the distance between her ideal point and the respective choice location---utility is maximized if an Actor can choose a policy located exactly on her ideal point, and utility is "lost" for choices farther and farther from her ideal point. The functional form of utility loss is an assumption made by the researcher---some scholars assume that utility loss follows a Gaussian curve, while others choose a quadratic utility loss [@clinton-jackman-rivers:2004:ideal]. For this analysis, we assume a quadratic utility loss.^[
  Researchers typically avoid linear losses for technical reasons: a linear utility loss function is non-differentiable at the ideal point because function comes to a point. This prevents the researcher from using differential calculus to find a point of maximum utility.
]
<!------- TO DO ---------
- differentiable cite?
------------------------->

The choice of quadratic loss implies a utility function over the *squared distance* between an Actor and a choice location. The utility Actor $i$ receives from choosing Left or Right are given by utility functions $U_{i}\left(L_{j}\right)$ and $U_{i}\left(R_{j}\right)$, respectively. With quadratic utility loss, these utility functions take the form
\begin{align}
  \begin{split}
    U_{i}\left(R_{j}\right) = - \left( \theta_{i} - R_{j}\right)^{2} + e^{\mathtt{R}}_{ij} \\ 
    U_{i}\left(L_{j}\right) = - \left( \theta_{i} - L_{j}\right)^{2} + e^{\mathtt{L}}_{ij},
  \end{split}
  (\#eq:utility-fns)
\end{align}
where $e^{\mathtt{R}}_{ij}$ and $e^{\mathtt{L}}_{ij}$ are the idiosyncratic error terms for the Right and Left alternatives, respectively. I sometimes refer to the quadratic utility loss as the "deterministic" component of the Actor's utility function, while the idiosyncratic error terms are "stochastic" components.^[
  Foreshadowing: Statistical estimates for model parameters are possible once we make distributional assumptions about the stochastic utility errors.
]



```{r plot-utility}
utility_plot +
  geom_line(
    data = utility,
    aes(x = x, y = utility_loss)
  ) +
  theme(legend.position = "none") +
  annotate(geom = "linerange",
    x = dd$Right, ymax = 0, ymin = -u_scale*(dd$Actor - dd$Right)^2,
    linetype = "dashed"
  ) +
  annotate(geom = "linerange",
    x = dd$Left, ymax = 0, ymin = -u_scale*(dd$Actor - dd$Left)^2,
    linetype = "dashed"
  ) +
  geom_point() +
  NULL
```

<!-- break -->

With these utility functions laid out, we can write Actor $i$'s decision as a comparison of the utilities received by choosing Right or Left. Let $y_{ij}$ indicate the Actor's choice of Right or Left, where Right is coded $1$, and Left is coded $0$. The model so far implies that $y_{ij} = 1$ (Actor chooses Right) if their utility is greater for Right than for Left.
\begin{align}
  y_{ij} = 1 &\iff U_{i}\left(R_{j}\right) > U_{i}\left(L_{j}\right)
  (\#eq:choice-iff)
\end{align}
To visualize this choice, I represent the deterministic components of Equation&nbsp;\@ref(eq:choice-iff) in Figure&nbsp;\@ref(fig:plot-utility), omitting the stochastic utility terms. The parabola represents $i$'s fixed utility loss for any choice along the ideological continuum, owed to her distance from that choice. The vertex of the parabola is at the Actor's location, indicating that she would maximize her spatial utility if she could choose a policy located exactly at her ideal point. Dashed lines below the Left and Right alternatives represent the utility loss owed to the Actor's distance from those specific choices. In the current example, the Actor is closer to Right than to Left, so she receives greater utility (or, less utility *loss*) by choosing Right instead of Left.

```{r plot-utility, include = TRUE, fig.height = 2, fig.width = 8, out.width = "100%", fig.cap = "A representation of quadratic utility loss over policy choices"}
```

It is important to remember that Figure&nbsp;\@ref(fig:plot-utility) shows only the deterministic component of choice task $j$; random error components $e^{\mathtt{R}}_{ij}$ and $e^{\mathtt{L}}_{ij}$ are omitted. With idiosyncratic utility error incorporated, Equation&nbsp;\@ref(eq:choice-iff) implies that even though the Actor's distance to Right is smaller than her distance to Left, there is still a probability that $i$ chooses Left. This probability depends on the instantiated values of the idiosyncratic error terms for each choice. These error terms represent the accumulation of several possible, non-ideological shocks to utility: random misperceptions about the policy locations, domain-specific considerations about a choice that aren't summarized by ideology, and so on. Supposing that these idiosyncratic terms follow some probability distribution, we can restate Equation&nbsp;\@ref(eq:choice-iff) with probabilistic notation:
\begin{align}
  \begin{split}
  \mathrm{Pr}\left(y_{ij} = 1\right) &= 
    \mathrm{Pr}\left(U_{i}\left(R_{j}\right) > U_{i}\left(L_{j}\right) \right) \\
  &=
    \mathrm{Pr}\left( 
      -\left( \theta_{i} - R_{j}\right)^{2} + e^{\mathtt{R}}_{ij} > 
      -\left( \theta_{i} - L_{j}\right)^{2} + e^{\mathtt{L}}_{ij}
    \right) \\
  &= \mathrm{Pr}\left(
      \left( \theta_{i} - L_{j}\right)^{2} -
            \left( \theta_{i} - R_{j}\right)^{2} 
        > e^{\mathtt{L}}_{ij} - e^{\mathtt{R}}_{ij}
      \right)
\end{split}
(\#eq:choice-prob)
\end{align}  
The intuition for Equation&nbsp;\@ref(eq:choice-prob) is that the Actor will choose the policy alternative that is nearest to her *unless* idiosyncratic (non-policy) factors overcome her ideological considerations. Supposing that the Actor is closer to Right than to Left, $\left( \theta_{i} - L_{j}\right)^{2}$ will be greater than $\left( \theta_{i} - R_{j}\right)^{2}$, capturing the Actor's deterministic inclination to prefer Right over Left. The only way for $i$ to choose Left would be if the idiosyncratic utility of Left over Right exceeded the Actor's deterministic inclinations.
  <!-- to do: assumes? -->
<!-- Note that the model so far assumes only that utility loss is quadratic over the ideological space and that the "coefficient" on the quadratic loss is equal for the distances to both Left and Right.  -->


Equation&nbsp;\@ref(eq:choice-prob) can be rearranged to reveal an appealing appealing functional form for $i$'s choice probability. First, expand the polynomial terms on the left side of the inequality...
\begin{align}
\begin{split}
  \mathrm{Pr}\left(y_{ij} = 1\right) &= 
    \mathrm{Pr}\left(
           \left( \theta_{i} - L_{j}\right)^{2} -
           \left( \theta_{i} - R_{j}\right)^{2} 
           > e^{\mathtt{L}}_{ij} - e^{\mathtt{R}}_{ij} 
         \right) \\
  &= \mathrm{Pr}\left(
      \theta_{i}^{2} - 2\theta_{i}L_{j} + L_{j}^{2} 
      - \theta_{i}^{2} + 2\theta_{i}R_{j} - R_{j}^{2}
      > e^{\mathtt{L}}_{ij} - e^{\mathtt{R}}_{ij} 
    \right) \\
  &= \mathrm{Pr}\left(
      2\theta_{i}R_{j} - 2\theta_{i}L_{j} + 
      L_{j}^{2} - R_{j}^{2}
      > e^{\mathtt{L}}_{ij} - e^{\mathtt{R}}_{ij} 
    \right)
\end{split}
(\#eq:expand-utility)
\end{align}

From here, there are two factorizations that reveal convenient expressions for important constructs in the model.
\begin{align}
\begin{split}
  \mathrm{Pr}\left(y_{ij} = 1\right) &=     
    \mathrm{Pr}\left(
      2\theta_{i}R_{j} - 2\theta_{i}L_{j} + 
      L_{j}^{2} - R_{j}^{2} 
      > e^{\mathtt{L}}_{ij} - e^{\mathtt{R}}_{ij}
    \right) \\
  &= 
    \mathrm{Pr}\left(
      2\theta_{i}R_{j} - 2\theta_{i}L_{j} + 
      \left(R_{j} - L_{j}\right)
      \left(R_{j} + L_{j}\right) 
      > e^{\mathtt{L}}_{ij} - e^{\mathtt{R}}_{ij} 
    \right) \\
  &= \mathrm{Pr}\left(
      2\left(R_{j} - L_{j}\right)
      \left(\theta_{i} - \frac{R_{j} + L_{j}}{2}\right)
      > e^{\mathtt{L}}_{ij} - e^{\mathtt{R}}_{ij} 
    \right)
\end{split}
(\#eq:factor-item-params)
\end{align}
The first manipulation is to decompose $L_{j}^{2} - R_{j}^{2}$ into the two factors $\left(R_{j} - L_{j}\right)\left(R_{j} + L_{j}\right)$. The second manipulation is to factor $2\left(R_{j} - L_{j}\right)$ out of the left-side of the inequality. We perform these manipulations because the resulting terms are appreciably more interpretable than before. First, note that $\frac{R_{j} + L_{j}}{2}$ is a formula for the midpoint between the Left and Right locations. This means that the expression $\theta_{i} - \frac{R_{j} + L_{j}}{2}$ intuitively conveys which policy alternative is closer to the Actor. If the Actor is closer to Right than to Left, $\theta_{i}$ will be greater than the midpoint, and vice versa if she were closer to Left. Second, the $2\left(R_{j} - L_{j}\right)$ term captures how far apart the policy alternatives are from one another, increasing as the distance between Right and Left increases. Together, the left side of the inequality succinctly describes the deterministic component of the Actor's ideological choice: is she closer to the Left or Right policy, and by how much?

The final manipulation is to simplify the terms above, which results in a convenient parameterization for statistical estimation.
\begin{align}
\begin{split}
  \mathrm{Pr}\left(y_{ij} = 1\right) 
  &= 
    \mathrm{Pr}\left(
      2\left(R_{j} - L_{j}\right)
      \left(\theta_{i} - \frac{R_{j} + L_{j}}{2}\right)
      > e^{\mathtt{L}}_{ij} - e^{\mathtt{R}}_{ij} 
    \right) \\
  &= \mathrm{Pr}\left(
      \beta_{j}\left(\theta_{i} - \kappa_{j}\right)
      > \epsilon_{ij}
    \right),
\end{split}
(\#eq:choice-2p)
\end{align}
This results in the "discrimination parameter" $\beta_{j} = 2\left(R_{j} - L_{j}\right)$, the "midpoint" or "cutpoint" parameter $\kappa_{j} = \dfrac{R_{j} + L_{j}}{2}$, and a joint error term $\epsilon_{ij} = e^{\mathtt{L}}_{ij} - e^{\mathtt{R}}_{ij}$.^[
  The names for these parameters are adapted from item-response theory (IRT), an area of psychometrics that is similarly interested in inferring latent traits from observed response data. I discuss the connection between this model and the IRT model in the next section.
]
Parameterizing the model in this way expresses the utility comparison in a simpler, linear form. Similar to Equation&nbsp;\@ref(eq:factor-item-params) above, $\theta_{i} - \kappa_{j}$ shows how far the Actor is from the midpoint between Left and Right, and $\beta_{j}$ captures behaves as a "slope" on this distance: the distance from the midpoint has a _stronger influence_ when the policy alternatives are farther from one another, since more utility is lost over larger spatial distances. I explore the intuitions of this functional form more thoroughly in the following section.

A complete statistical model is obtained by making a parametric assumption for the distribution of $\epsilon_{ij}$. Assuming that $\epsilon_{ij}$ is a draw from a standard Normal distribution,^[
  This implies that $\mathrm{E}\left(e^{\mathtt{L}}_{ij}\right) = \mathrm{E}\left(e^{\mathtt{R}}_{ij}\right)$ and that $\mathrm{Var}\left(e^{\mathtt{L}}_{ij} - e^{\mathtt{R}}_{ij}\right) = s_{j} = 1$. 
  For a given choice $j$, imposing a scale restriction on the magnitude of $s_{j}$ is not problematic since the ideological scale is latent and can be arbitrarily stretched. The important assumption is that the scale of the errors for each choice are equal across individuals: $s_{ij} = s_{j}$ for all $i$. 
]
Equation&nbsp;\@ref(eq:choice-2p) implies a probit regression model for the probability that Actor $i$ chooses Right on choice $j$:
\begin{align}
\begin{split}
  \mathrm{Pr}\left(y_{ij} = 1\right) &= 
  \mathrm{Pr}\left(
      \beta_{j}\left(\theta_{i} - \kappa_{j}\right)
      > \epsilon_{ij}
    \right) \\
  &= 
  \mathrm{Pr}\left(
    \beta_{j}\left(\theta_{i} - \kappa_{j}\right) - \epsilon_{ij} > 0 
  \right) \\
  &= 
  \Phi\left( \beta_{j}\left(\theta_{i} - \kappa_{j}\right) \right),
\end{split}
(\#eq:logit-irt)
\end{align}
where $\Phi(\cdot)$ is the cumulative Normal distribution function. Many IRT models assume that $\epsilon_{ij}$ follows a standard Logistic distribution, [for example @londregan:1999:ideal-pts], resulting in a logistic regression model rather than a probit model.^[
  A technical point of difference between the probit and logit model is the way parameters are scaled to yield the final line of Equation&nbsp;\@ref(eq:choice-2p). If it is assumed that $\epsilon_{ij}$ is a Logistic draw with scale $s_j$, this implies that 
  $\mathrm{Var}\left(e^{\mathtt{L}}_{ij} - e^{\mathtt{R}}_{ij}\right) = \dfrac{s_{j}^{2}\pi^{2}}{3}$, where it is assumed that $s_{j} = 1$ for the standard Logistic model. 
]
As I show below, the probit model facilitates the group-level model much more easily than the logit model.

<!------- TO DO ---------
- Does londregan prefer logit due to some identifiability concern???
- cite CW/Fox
------------------------->



### The "Item Response Theory" Approach to Survey Response

Past ideal point modeling in political science has noted the similarity between the logit and probit models and models developed under item response theory (IRT) in psychometrics [for example, @londregan:1999:ideal-pts]. The mission of IRT models in psychometrics is similar to ideal point models in political science: measuring latent features in the data given individuals' response patterns to various stimuli. The canonical example is in education testing, where tests are used to measure a student's unobserved academic "ability" level. This section elaborates on important theoretical and mathematical features of IRT models by drawing an analogy between the educational testing context to the current setting of survey response. 

##### Latent traits
The first important feature to note about IRT models is that they are *measurement models*. The goal of a measurement model is to use observed data $\mathbf{y}$ to estimate some construct of theoretical interest ${\theta}$, supposing that there is a distinction between the two. The observed data $\mathbf{y}$ may be affected by ${\theta}$, but there is no guarantee of a one-to-one correspondence between the two because $\theta$ is not directly observed. We might think of a measurement model represented with general notation $\mathbf{y} = f({\theta}, {\sigma})$, where ${\sigma}$ represents some vector of auxiliary model parameters.

In an educational testing context, students are taking standardized tests intended to measure their academic "ability" levels. An analyst scoring the tests cannot observe a student's ability directly---it isn't clear what that would even mean. They do, however, observe the student's answers to known test questions. IRT models provides a structure to infer abilities from the student's pattern of test answers. The context of policy choice is similar. We cannot observe any individual's political ideology directly, but we theorize that it affects their responses to survey items about policy choices. The IRT setup lets us summarize an individual's policy preferences by analyzing the structure of their responses to various policy choices.

<!------- TO DO ---------
- Literature about "ideology" (Clinton piece?)
- "traits"
------------------------->

##### Item parameters

One parameter

Two parameter

Zero parameter


Contours:

- measurement model;
    - Latent construct 
    - Elicited responses aren't perfectly informative: error, question bias
- items $j$, what do we learn about them
- Rasch models



### Ability-only model

We begin with a simple model where an individual $i$ gives a conservative response to item $j$ if they derive more utility than they would from a liberal response. Their utility function is made up of their ideal point $\theta_{i}$ and idiosyncratic error $\varepsilon_{\mathit{ij}}$. We can assume that the threshold between liberal and conservative responses is at $0$; this is merely a scale and location restriction on what is otherwise an unstructured ideological space. 
\begin{align}
  y_{\mathit{ij}} &= I\left( \theta_{i} + \varepsilon_{\mathit{ij}} > 0 \right)
  (\#eq:ability-only)
\end{align}
This means $y_{\mathit{ij}} = 1$ when the utility is positive, and $y_{\mathit{ij}} = 0$ otherwise.

Assuming that the idiosyncratic utility is $\varepsilon_{\mathit{ij}} \sim \mathrm{Normal}(0, 1)$ gives us a probit model,
\begin{align}
  \mathrm{Pr}\left( y_{\mathit{ij}} = 1 \right) 
    &= 
  \Phi \left( \theta_{i} \right),
  (\#eq:ability-only-probit)
\end{align}
where $\Phi(\cdot)$ is the Normal cumulative distribution function. This assumption restricts $\theta$ such that the probability of a correct response is $50\%$ when $\theta = 0$. The scale of $\theta$ is also restricted by the assumption that $\mathrm{Var}\left[\varepsilon_{\mathit{ij}}\right] = 1$.

```{r plot-ability-model, include = TRUE, out.width = "80%", fig.width = 6, fig.height = 4, fig.cap = "Ability-only model of item response"}
# No systematic item effects, only a normal CDF
ability_bound <- 3

tibble(theta = seq(-ability_bound, ability_bound, .01)) %>%
  mutate(p = pnorm(q = theta, mean = 0, sd = 1)) %>%
  ggplot(aes(x = theta, y = p)) +
    coord_cartesian(xlim = c(-ability_bound, ability_bound),
                    ylim = c(0, 1)) +
    geom_segment(aes(x = -4, xend = 0, y = 0.5, yend = 0.5), 
                 linetype = "dotted", size = 0.25, color = "steelblue") +
    geom_segment(aes(x = 0, xend = 0, y = 0.5, yend = -1), 
                 linetype = "dotted", size = 0.25, color = "steelblue") +
    geom_line() +
    annotate(geom = "text", x = 0, y = 0.5, 
             label = TeX("$\\mathrm{Pr}(\\mathit{y}_{\\mathit{ij}} = 1) = 0.5$\n when $\\theta_{\\mathit{i}}$ is $0$"),
             hjust = -0.2,
             family = "Minion Pro") +
    labs(x = TeX("Ideal Point $(\\theta_{\\mathit{i}})$"),
         y = "Probability of Conservative Response") 
```


### One-parameter Rasch model

Ability model assumes no "item effects"---there is no systematic variation at the item level that would lean to a correlated effect for one item across individuals.
  <!-- to do: e.g. -->
  <!-- such as ... difficulty of a test question -->

This is the justification for the one-parameter Rasch model, developed in an educational testing framework. The model defines the probability that individual $i$ answers item $j$ "correctly" as a comparison between an individual $i$'s latent "ability" ($\theta_{i}$) and a test question $j$'s "difficulty" ($\kappa_{j}$). Assuming that other disturbances affecting item responses ($\varepsilon_{\mathit{ij}}$) are normally distributed, individual $i$'s response to item $j$ is
\begin{align}
  y_{\mathit{ij}} &= 
    I\left( \left[\theta_{i} - \kappa_{j}\right] + \varepsilon_{\mathit{ij}} > 0 \right),
    (\#eq:onepar-model)
\end{align}
which yields a probit model for a discrete outcome.
\begin{align}
  \mathrm{Pr}\left( y_{\mathit{ij}} = 1 \right) 
  &= 
  \Phi \left( \theta_{i} - \kappa_{j} \right)
  (\#eq:onepar-probit)
\end{align}

Adapting this model to the context of survey response, an individual respondent $i$ is asked to choose between two alternatives on policy item $j$. We let $\theta_{i}$ be $i$'s ideal point on a liberal-conservative dimension, where larger values of $\theta$ are more conservative, and $\kappa_{j}$ is the midpoint between the liberal and conservative policy alternatives. Respondent $i$ is more likely to give a conservative response on policy $j$ if their ideal point is to the "right" of the item midpoint (their ideal point is nearer to the conservative policy choice), but their response is affected by other normally distributed factors and is thus predicted only probabilistically.


### Two-parameter Rasch Model

The item-response model makes sense only under the assumption that policy issues are choices along a single underlying dimension. The one parameter model allows non-ideological factors to affect item responses but assumes that these extraneous factors are independent across items and individuals. Policy issues don't all have the same salience to partisan and ideological debate, however, so political scientists often model policy item responses using a two-parameter model. The two-parameter model includes the additional item parameter $\beta_{j}$, which relaxes the assumption that all items are equally related to latent ideology. This "discrimination" parameter captures how strongly the item divides the responses of liberals and conservatives. 
\begin{align}
  \mathrm{Pr}\left( y_{\mathit{ij}} = 1 \right)
  &= 
  \Phi \left( \beta_{j}\left[\theta_{i} - \kappa_{j}\right] \right),
  (\#eq:twopar-model)
\end{align}


```{r plot-example-iccs, include = TRUE, echo = FALSE, fig.cap = str_glue("Examples of item characteristic curves for {ex_items} simulated items."), fig.width = 6, fig.height = 4, out.width = "80%"}
# ---- create example ICCs -----------------------

set.seed(2012)

ex_items <- 5

ex_item_params <- 
  tibble(cutpoint = rnorm(ex_items, 0, 1.5), 
         discrimination = rlnorm(ex_items, 0, 1)) %>%
  mutate(item = 1:n()) 

ex_iccs <- ex_item_params %>%
crossing(theta = seq(-5, 5, .01)) %>%
  mutate(eta = discrimination * (theta - cutpoint),
         prob = pnorm(eta))


# ---- plot them -----------------------

ggplot(ex_iccs, aes(x = theta, y = prob)) +
  geom_point(data = ex_item_params, aes(x = cutpoint, y = 0.5),
             size = 1.5) +
  geom_line(aes(group = as.factor(item)), show.legend = FALSE) +
  scale_y_continuous(breaks = c(0, .5, 1)) +
  scale_x_continuous(breaks = c(-3, 0, 3),
                     labels = c("Liberal", "Moderate", "Conservative")) +
  scale_color_brewer(palette = "Set2") +
  labs(x = TeX("Ideal Point $(\\theta_{\\mathit{i}})$"), y = "Probability of Conservative Response") +
  NULL
```

```{r}
knitr::knit_exit()
```

<!-- IRT model parameters -->

How do we interpret the parameters $\beta_{j}$ and $\alpha_{j}$? The discrimination parameter behaves as a "coefficient" for Actor $i$'s ideal point in Equation&nbsp;\@ref(eq:glm), meaning that the Actor's choice is more or less responsive to her policy preferences as $\beta_{j}$ varies. This happens as a function of the ideological distance between the Right and Left alternatives; holding the error variance ($\sigma_{j}$) constant, the discrimination parameter $\beta_{j} = \dfrac{1}{\sigma_{ij}}2(\rho_{j} - \lambda_{j})$ grows in magnitude Right and Left alternatives grow farther apart. The difficulty parameter $\alpha_{j} = \dfrac{1}{\sigma_{ij}}\left(\lambda_{j}^{2} - \rho_{j}^{2}\right)$ serves as an intercept, capturing the phenomena that some policy choices present alternatives that are jointly more conservative or liberal on average. It is understand $\alpha_{j}$'s behavior by considering special cases first. Suppose that $\alpha_{i} = 0$, which occurs if the Right and Left locations are equidistant from zero in opposite directions. In such a case, an Actor whose ideal point is located exactly at $\theta_{i} = 0$ would be indifferent (in expectation) to the choice of Left or Right, and the probability of choosing Right would be $0.5$.^[
  This holds in logit and probit models, since $\mathrm{logit}^{-1}(0) = 0.5$ and $\Phi(0) = 0.5$.
]
If we move the Right alternative in a conservative direction (increasing the value of $\rho_{j}$), the result is a decrease in $\alpha_{j}$. This implies a lower probability that $i$ chooses Right, all else equal, because the Right alternative is moving farther from an Actor located at $\theta_{i} = 0$. The opposite intuition holds as the Left position becomes increasingly progressive, resulting in larger values of $\alpha_{j}$ that imply a higher probability of choosing Right, all else equal.^[
  In a special case that Right and Left alternatives are located in exactly the same location, the result is $\alpha_{j} = \beta_{j} = 0$, leading all Actors to choose Right with probability $0.5$. This result represents a situation where policy preferences are not systematically related to the choice whatsoever, and only idiosyncratic error affects the choice of Right or Left. Although the model implies that this result is *mathematically* possible, it is not realistic to expect any of the policy choices in this dissertation to induce this behavior.
]

Others have noted [see @londregan:1999:ideal-pts] that the reduced form is equivalent to the "two-parameter Rasch model" from the psychometric field of item-response theory (IRT). I briefly discuss the intuition of these models and how their intuitions relate to ideal point models.



<!-- RASCH MODEL
^[
  The names "difficulty" and "discrimination" come from models developed in an education testing, where some test questions may differ in their difficulty as well as how well the question discriminates between high- and low-ability test takers. See e.g.\ @fox:2010:bayesian-IRT.
] -->




<!-- parametric assumptions -->
<!-- Up to this point, we have assumed symmetric and equally-weighted quadratic utility loss over ideological distances. No parametric assumptions have yet been made about the idiosyncratic errors or the difference in errors.  -->



This would imply that $\alpha_{j} = 0$, 

 increase (that is, move to the right). 

An intuition for this parameter is that


 for the is that $i$'s choice places greater weight 





<!-- to do: error -->
<!-- Assumptions about the error? Interpretation of the "random" vs "unsystematic" meaning of error? -->


<!-- What is the formal model intuition?  -->

<!-- ### Unobserved Utility and Observed Choices -->

Ideology itself is unobservable. It affects the way people feel about policy, but we don't directly observe it. Explicit measures of ideological self identification are not themselves reliable because people don't understand those items
  <!-- to do: TK -->
  <!-- cite -->
and because they aren't great at predicting policy views either
  <!-- to do: TK -->
  <!-- cite -->


While we do not directly observe ideal points, the individuals make are observable functions of unobserved preferences. These problems are common in micro-economic modeling---what inferences can be made about unobserved utilities from observed choices? Ideal point models, as a result, typically have a theoretical basis in a utility model. 


Measurement model: 

- We can't see the construct that we want, so we estimate it
- by building a model that relates the unobserved construct to observed data

<!-- Utility framework

- normal error
- symmetric utility loss on either end of kappa

-->







### Applications of the IRT Approach

<!-- ### Estimating Citizen Ideology with Survey Data -->

<!-- ### Statistical Estimation of Ideal Points -->

<!-- Relation to additive indices (ability-only models) -->



## Modeling Party-Public Ideology in Congressional Districts

This section outlines my group-level ideal point model for party publics. It begins by describing the connection between the individual-level IRT model and the group-level model and its implication for the parameterization of the model (Section&nbsp;\@ref(sec:how-to-group)). I then lay out the hierarchical model for party-public ideal points in its static form (Section&nbsp;\@ref(sec:geographic-model)) and its dynamic form (Section&nbsp;\@ref(sec:dynamic-model)). Lastly, I discuss more technical features of model implementation, including prior distributions and model identification (Section&nbsp;\@ref(sec:priors-identification)) and parameterization in Stan (Section&nbsp;\@ref(sec:stan-setup)).


<!-- ### Static Model -->

### Group-Level IRT Setup {#sec:how-to-group}

<!--  -->


To motivate the intuition of the group model, we begin the intuition of the individual and then reparameterize it.

Assume that individuals answer policy items according to their own individual ideal points. We observe a binary response from individual $i$ to item $j$, which we model as a probabilistic outcome with probability $\pi_{\mathit{ij}}$.
\begin{align}
  y_{\mathit{ij}} &\sim \mathrm{Bernoulli}(\pi_{\mathit{ij}})
  (\#eq:y-i-response)
\end{align}
Our model for the response probability is given by a probit model, following the utility intuition described above.
\begin{align}
  \pi_{\mathit{ij}} 
  &= 
  \Phi\left( \beta_{j}\left[ \theta_{i} - \kappa_{j} \right] \right)
  (\#eq:y-i-probit)
\end{align}

Let $\sigma_{j} = \beta^{-1}_{j}$, so item discrimination is instead expressed as a "dispersion" parameter [@fox:2010:bayesian-IRT; @caughey-warshaw:2015:DGIRT].
\begin{align}
  \pi_{\mathit{ij}} &= \Phi\left( \frac{\theta_{i} - \kappa_{j}}{\sigma_{j}} \right)
  (\#eq:y-i-irt)
\end{align}

Assume that individuals are normal draws from the mean of their group, where a group is a party in a district.^[
  Notation for Normal distributions will always describe the scale parameter in terms of standard deviation $\sigma$ instead of variance $\sigma^2$. This keeps the notation consistent with the way Normal distributions are expressed in Stan code.
]
\begin{align}
  \theta_{i} &\sim \mathrm{Normal}\left( \theta_{g[i]}, \sigma_{g[i]} \right)
  (\#eq:theta-i-draw)
\end{align}

The outcome data at the group level, rather than individual Bernoulli outcomes, are expressed as grouped Binomial outcomes $Y_{gj}$: the number of conservative responses in group $g$ to item $j$ given the total number of responses per group per item, $n_{gj}$. The probability that a randomly selected individual gives a conservative response (the "true conservative probability" to item $j$ in group $g$) is $\bar{\pi}_{gj}$.
\begin{align}
  Y_{gj} &= \mathrm{Binomial}\left( \bar{\pi}_{gj}, n_{gj} \right)
  (\#eq:y-g-binom)
\end{align}

The probability for each item-group is given a probit model from the spatial utility model.
\begin{align}
  \bar{\pi}_{gj} 
  &= 
  \Phi\left(
    \frac{\theta_{g} - \kappa_{j}}{
          \sqrt{ \sigma_{g}^{2} + \sigma^{2}_{j}}}
  \right),
  (\#eq:y-g-probit)
\end{align}
where $\sigma_{g}$ is the standard deviation of ideal points within group $g$, which is introduced because we are now estimating the mean response within a group rather than an individual item response. Larger values of $\sigma_{g}$ indicate more uncertainty over the item response and attenuate $\bar{\pi}_{gj}$ toward $50\%$.^[
  The binomial setup assumes that each of the $n_{gj}$ trials is independent conditional on the mean ideal point $\theta_{g}$ and the item parameters $\kappa_{j}$ and $\sigma_{j}$. This assumption is violated if individuals in a group answer multiple items---errors are not independent across items. I describe a design-weighting correction for this assumption in Section&nbsp;\@ref(sec:stan-setup).
]
<!-- to do: section -->
  <!-- That's not the right section? -->


### Geographic Model for Group Means and Scales {#sec:geographic-model}

Estimates for $\theta_{g}$ and scales $\sigma_{g}$ are improved using a hierarchical model.
  <!-- to do: scales -->
<!-- why "scales" -->


- Using geographic data to improve estimation
- Hierarchical model accounts for multiple sources of variation
- Partial pooling estimates for groups without as much data
- First, using traditional notation. Reparameterization in Section&nbsp;\@ref(sec:stan-setup) greatly improves the estimation in Stan.

<!-- ##### Conventional hierarchical notation -->
We consider it a draw a distribution with hypermean $\bar{\theta}_{g}$ and scalar standard deviation $\psi_{\theta}$
\begin{align}
  \theta_{g} &\sim \mathrm{Normal}\left(\bar{\theta}_{g}, \psi_{\theta}\right)
  (\#eq:theta-g-draw)
\end{align}
The hierarchical setup improves estimates by casting the hypermean as a conditional mean from a regression on group data. We specify this regression using geographic-level data from the districts and states where each group is located:
\begin{align}
  \bar{\theta}_{g} &= \beta_{0p[g]} + X_{d[g]}\beta_{p[g]} + \alpha_{s[g]p[g]},
  (\#eq:theta-g-hlm)
\end{align}
where $\beta_{0}$ is a constant, $X_{d}$ represents district-level data with coefficients $\beta$, and $\alpha_{s}$ represents state effects. A key feature of the hierarchical model is that the parameters are indexed by $p$ and thus dependent on the party to which $g$ belongs. This means there are two constants $\beta_{0p}$  where $p \in \{1, 2\}$ indexes party. Group-level covariates $X_{g}$ have coefficient vectors $\beta_{p}$ that vary by party as well. We include this flexibility because geographic covariates (such as racial composition) may have different correlations with ideal points depending on the party. This is a departure from the model laid out by @caughey-warshaw:2015:DGIRT, which holds the geographic regression fixed for all groups in the data.

We use a similar hierarchical regression for group scales (suppressing the $g$ subscript which is implied by the combination of $d$ and $p$).
\begin{align}
  \log \left( \sigma_{g} \right) 
  &\sim 
  \mathrm{Normal}\left( 
    \delta_{0p} + X_{d}\delta_{p} + \eta_{sp}, 
    \psi_{\sigma}
  \right),
  (\#eq:sigma-g-draw)
\end{align}
where $\delta_{0p}$ represents constants for each party, $\delta_{p}$ is a party-specific vector of coefficients on district features, and $\eta_{sp}$ are party-specific state effects.

The state effects are in turn regressions on state features.
\begin{align}
\begin{split}
  \alpha_{sp} &\sim
    \mathrm{Normal}\left(Z_{s}\gamma_{p} + \rho_{r[s]p}, \psi_{\alpha}\right), \\
  \eta_{sp} &\sim 
    \mathrm{Normal}\left(Z_{s}\zeta_{p} + \lambda_{r[s]p}, \psi_{\eta}\right),
\end{split}
(\#eq:state-effects-draws)
\end{align}
where $Z_{s}$ contains state covariates which have party-specific coefficients $\gamma_{p}$ (for group means) or $\zeta_{p}$ (for group scales). Each state effect is a function of a party-specific region effect $\rho_{rp}$ (for group means) and $\lambda_{rp}$ (for group scales) for Census regions indexed $r$.

### Weighting for Sample Design and Repeated Observations per Individual {#sec:model-weights}

### Dynamic Model {#sec:dynamic-model}

### Priors and Identification {#sec:priors-identification}





### Model Implementation in Stan {#sec:stan-setup}

<!-- Non-Centered Parameterization -->
Hierarchical models often have posterior distributions whose curvature presents difficulties for sampling algorithms [@betancourt:2015:hamiltonian; @papaspiliopoulos-et-al:2007:parameterizations]. To improve the estimation in Stan,
  <!-- to do: Stan -->
<!-- Flag early on that I'll be setting things up to make things easier in Stan -->
I parameterize the hierarchical models in the "non-centered" rather than the "centered" form. Whereas the centered form considers $\theta_g$ as a random draw from a distribution, the non-centered parameterization considers $\theta_{g}$ as a function of the hypermean and a random error.
\begin{align}
  \theta_{g} &=
    \beta_{0p} + X_{g}\beta_{p} + \alpha_{s[g]} + z_{g}\varepsilon
  (\#eq:theta-g-noncenter)
\end{align}
where $z_{g}\varepsilon$ represents a group-level error term. It is composed of a $z$-score that is $\mathrm{Normal}\left(0, 1\right)$ and a scale parameter $\varepsilon$. The non-centered paramaterization has the same algebraic behavior as the centered parameterization, but it has the practical effect of improving Monte Carlo sampling by de-correlating the parameters in the hierarchical model. We can also de-center the $\alpha_{s[g]}$ term
\begin{align}
  \alpha_{s[g]p} &= Z_{s[g]}\gamma_{p} + u_{sp}\tau
  (\#eq:state-effect-noncenter)
\end{align}
where $u_{sp}$ is a $z$-score distributed $\mathrm{Normal}\left(0, 1\right)$, and $\tau$ is a scale factor.

<!-- Heteroskedastic model -->
We also have a hierarchical model that predicts the ideal point standard deviation within each group, $\sigma_{g}$. This makes the model "heteroskedastic"---we are modeling the mean ideal point within each group and the variance, conditional on hierarchical covariates. The model for $\sigma_{g}$ in non-centered form is as follows:
\begin{align}
  \log(\sigma_{g}) &= X_{g}\delta_{p} + Z_{s[g]}\eta_{p} + m_{g}\nu + m_{sp}\lambda
  (\#eq:sigma-g-hetero-noncenter)
\end{align}
Where $X_{g}$ and $Z_{sp}$ are the same group- and state-level covariates as the above regression, $\delta_{p}$ and $\eta_{p}$ are party-varying coefficients. The terms $m_{g}\nu$ and $m_{sp}\lambda$ are "factored" error terms for groups and states, where $m_{g}$ and $m_{sp}$ are each distributed $\mathrm{Normal\left(0, 1\right)}$, while $\nu$ and $\lambda$ are scale factors.
  <!-- to do: Factoring -->
<!-- Describe it as factoring the scale parameter out of an error term} -->


##### IRT Model

##### Noncentered Hierarchical Model

##### Any other vector/matrix tricks?

##### Read stan files into an appendix?

