# Bayesian Approach to Causal Inference in Political Science {#ch:causality}


I'm new!


```{r knitr-03-causality, include = FALSE, cache = FALSE}
source(here::here("assets-bookdown", "knitr-helpers.R"))
```

```{r r-03-causality, cache = FALSE}
library("knitr")
library("here")
library("magrittr")
library("tidyverse")
library("extrafont")
library("latex2exp")
library("scales")
library("patchwork")
```


## General Orientation Toward Causal Inference

Taboo against explicit causal inference, (Grosz, Rohrer, Thoemmes)

Whatever the Hernán (2018) article says

The point of causal inference isn't to shame observational studies away from making causal claims. The point is to clarify which designs and assumptions enable certain causal claims and which do not.

Strictly speaking, causality is never observed. It is only inferred with the help of assumptions.


## The Meaning of Parameters in Design and Estimation

What was I trying to say here? 

The causal parameter is a feature of the _causal model_. The causal model is the model if you knew everything. When we impose assumptions for estimating, we can say under which conditions this parameter can be _estimated_. The estimator of the parameter is a different thing. 

Bayesian estimation is a different _estimation_ framework given the same causal model. This is easy to see when we think about all Bayesian models (at least, MLE models) as special cases of Bayesian models with all log-likelihoods equally weighted. 

## Rubin's Vision: Posterior Distributions for Missing Potential Outcomes

## Parameterization and other Opportunities

Gelman "Parameterization" paper: parameterization for convenience often reveals a new family of models. For Bayesian inference, it reveals a new family of priors. For _causal_ inference, we should pay careful attention to the way parameterization bears on the priors that we do and do not feel comfortable using, and the way that parameterization presents opportunities to improve on other work.

```{r diff-means-example}
means_data <- 
  tibble(
    x = seq(-1, 1, .05),
    unif = ifelse(x >= 0, dunif(x, 0, 1), NA),
    triangle = 
      ifelse(
        x < 0, 
        (2 * (x - -1)) / ((1 - -1) * (0 - -1)), 
        (2 * (1 - x)) / ((1 - -1) * (1 - 0))
      )
  ) 

plot_prior <- function(data, x, y) {
  ggplot(data) +
  aes(x = {{x}}, y = {{y}}) +
  geom_ribbon(
    aes(ymin = 0, ymax = {{y}}),
    fill = primary_light
  ) +
  geom_line() +
  scale_y_continuous(limits = c(0, 1.5))
}
```

```{r plot-diff-means-example}
plot_prior(data = means_data, x = x, y = unif) +
  coord_cartesian(xlim = c(0, 1)) +
  labs(
    x = TeX("Group Mean: $\\mu_{z}$"),
    y = "Prior Density",
    title = "Prior Densities for Difference in Means",
    subtitle = "If means have Uniform(0, 1) priors"
  ) +
plot_prior(data = means_data, x = x, y = triangle) +
  labs(x = TeX("Treatment Effect: $\\mu_{1} - \\mu_{0}$")) +
  theme(
    axis.ticks.y = element_blank(),
    axis.text.y = element_blank(),
    axis.title.y = element_blank()
  ) 
```

For instance, consider a simple experiment with a binary outcome variable $y_{i}$ and binary treatment assignment $z_{i} \in \{0, 1\}$. Suppose that the treatment effect of interest is a difference-in-means, $\bar{y}_{z = 1} - \bar{y}_{z = 0}$, estimated from a linear probability model. This linear probability model might be parameterized in two ways. First is a conventional regression setup,
\begin{align}
  y_{i} &= \alpha + \beta z_{i} + \epsilon_{i}
  (\#eq:diff-means-example)
\end{align}
where $\alpha$ is the control group mean, $\alpha + \beta$ is the treatment group mean, $\beta$ represents the difference in means, and $\epsilon_{i}$ is a symmetric error term for unit $i$. With the model parameterized in this way, the researcher must specify priors for $\alpha$ and $\beta$. Suppose that the researcher gives $\beta$ a flat prior to represent ignorance about the treatment effect. An equivalent _likelihood model_ for the data would be to treat each observation as a function of its group mean $\mu_{z}$.
\begin{align}
  y_{i} &= z_{i}\mu_{1} + \left(1 - z_{i}\right)\mu_{0}  + \epsilon_{i}
  (\#eq:separate-means-example)
\end{align}
Although the treatment effect $\beta$ from Equation&nbsp;\@ref(eq:diff-means-example) is equivalent to the difference in means $\mu_{1} - \mu_{0}$ from Equation&nbsp;\@ref(eq:separate-means-example), the parameterization of the model affects the implied prior for the difference in means. If the researcher gives a flat prior to both $\mu_{\mathbf{z}}$ terms, the implied prior for the difference in means will not be flat. Instead, it will be triangular, as shown in Figure&nbsp;\@ref(fig:plot-diff-means-example). The underlying mechanics of this problem are well-known in applied statistics---if we continue adding parameters, the Central Limit Theorem describes how the resulting distribution will converge to Normality---but it takes the explicit specification of priors to shine a light on the consequences of default prior choices in a particular case. In particular it shows how even flat priors, which are popularly regarded as "agnostic" priors because of their implicit connection to maximum likelihood estimators, do not necessarily imply flat priors about the researcher's key quantities of interest. Rather, flat priors can create a variety of unintended prior distributions that do not match the researcher's expectations. I return to this important idea in the discussion about setting priors for a probit model in Section&nbsp;\@ref(sec:probit-prior).

```{r plot-diff-means-example, include = TRUE, fig.height = 4, fig.width = 8, out.width = "100%", fig.cap = "Model parameterization affects prior distributions for quantities of interest that are functions of multiple parameters or transformed parameters. The left panel shows that the difference between two means does not have a flat prior if the two means are given flat priors. Note that the $x$-axes are not fixed across panels."}
```

- equivalent parameterizations


## Example Case: Regression Discontinuity with a Binary Outcome

## Example Case: Meta-Analysis of Nonparametric Treatment Effects

