# Candidate Positioning in Primary Elections {#ch:positioning}


```{r knitr-04-1-positioning, include = FALSE, cache = FALSE}
source(here::here("assets-bookdown", "knitr-helpers.R"))
```


```{r packages-04-positioning, cache = FALSE}
library("knitr")
library("here")
library("magrittr")
library("tidyverse")
library("latex2exp")
library("scales")
# library("patchwork")
# library("ggforce")
library("broom")
library("tidybayes")
library("boxr"); box_auth()
library("ggdag"); 
theme_dag <- theme_dag(base_family = font_fam)
```

## Candidate Positioning in Primary Elections




## The Direct Effect of Partisan Preferences

### Causal Graphs

Collider bias: ideological proximity and quality both increase chances of victory. Conditioning on losers, an ideologically proximate losing candidate is more likely to be low-quality, and a high-quality losing candidate is more likely to be ideologically distant

```{r problem-dag}
problem_dag <- 
  dagify(
    C ~ T + P + U,
    P ~ T + U,
    exposure = "T",
    outcome = "C",
    coords = tribble(
      ~ name,      ~ x,    ~ y,
      "T",      0,      0,
      "C",      2,      0,
      "P",      1,      1,
      "U",      2,      1
    ),
    labels = c(
      "T" = "Partisan Ideology",
      "C" = "Candidate Positioning",
      "P" = "District Partisanship"
    )
  ) %>%
  tidy_dagitty() %>%
  # node_parents("C") %>%
  mutate(
    pt_label = case_when(
      name == "T" ~ "theta[pd]",
      name == "C" ~ "Y[cpd]",
      name == "P" ~ "P[d]",
      name == "U" ~ "U"
    )
  ) %>%
  print()
```


```{r plot-problem-dag}
ggplot(problem_dag) +
  aes(x = x, y = y, xend = xend, yend = yend) +
  geom_dag_edges() +
  geom_dag_node(
    aes(color = (name == "U"))
  ) +
  geom_dag_text(
    aes(label = pt_label), 
    parse = TRUE, 
    family = font_fam
  ) +
  scale_color_manual(values = c("TRUE" = "gray50", "FALSE" = primary)) +
  annotate(
    geom = "text",
    label = "District Partisanship (P) affects\nCandidate Positioning (C) but is also\naffected by District-Party Ideology (θ)",
    x = 0.2, y = 1,
    size = 3
  ) +
  theme_dag(
    legend.position = "none"
  ) + 
  expand_plot(
    expand_x = expand_scale(c(0.2, 0.2)), 
    expand_y = expand_scale(c(0.2, 0.2))
  ) +
  NULL  
```


```{r plot-problem-dag, include = TRUE, fig.width = 5, fig.height = 3, out.width = "80%", fig.cap = "A DAG that presents district partisanship as a collider along the path from district-party preferences to candidate positioning. Controlling for district partisanship can bias the causal estimate of district-party preferences in several ways. If there is a path $\\theta \\rightarrow P \\rightarrow Y$, then the effect of $\\theta$ does not represent the total effect. In the presence of intermediate confounder $U$, controlling for district partisanship induces collider bias by unblocking the path $\\theta \\leftarrow U \\rightarrow Y$."}
```

### Modeling Direct Effects

Controlled direct effect (intervene on ideology, intervene on fixed level of vote in the district):

Blackwell notation: $CDE_{i}(a, a', m) = Y_{i}(a, m) - Y_{i}(a', m)$.

More general _do_ notation: $p(CDE(a, a', m)) = p(y \mid do(A = a'))$ (not finished)

$p(y(\mathrm{do}(\theta), \mathrm{do}(v)) - y(\mathrm{do}(\theta'), \mathrm{do}(v)))$

Natural direct effect (intervene on ideology, fixing the _given_ level of vote in the direct):
$p\left( y\left(\mathrm{do}\left(\theta\right), v\right) - y\left(\mathrm{do}\left(\theta'\right), v\right) \right)$

Sequential-$g$ estimator: Joffe and Greene 2009; Vansteelandt 2009

### Bayesian Considerations

We observe $y_{g}\left(\theta_{g}\right)$, but we don't know $\theta_{g}$ with certainty. Instead let $p\left(\theta_{g}\right)$ be the posterior distribution of $g$'s average policy ideology as estimated from the group IRT model in Chapter&nbsp;\@ref(ch:model). Our observed outcome is a function of this unknown value, which we represent as $y_{g}\left(p\left(\theta_{g}\right)\right)$. 

Thus the causal effect of setting $\theta = \theta'$ for unit $g$ is
\begin{align}
  \tau_g &= y_{g}\left(\theta'\right) - y_{g}\left(p\left(\theta_{g}\right)\right),
\end{align}
or, written as a probability distribution 
\begin{align}
  p\left(\tau_{g}\right) &= p\left(y_{g}\left(\theta'\right) - y_{g}\left(\theta_{g}\right)\right)
\end{align}



## Empirical Approach


### Data

U.S. House primary data from @pettigrew-primary-data. 

### Model



