# Colophon

This project is open source and managed with Git.
A remote copy of the repository is available at <https://github.com/mikedecr/dissertation>.
Currently, the repository is at the following commit:

```{r echo = FALSE, include = TRUE}
# I've set eval=FALSE to ensure Travis-CI can run
# if you're not using Travis-CI, then eval=TRUE will be fine
library(git2r)
if (in_repository()) {
  summary(commits()[[1]]) 
} else { 
  message("We are not in a git repository") 
}
```

This version of the document was generated on `r Sys.time()`. 

All Bayesian models were estimated using the probabilistic programming language Stan.
Front-end interface to Stan and other data management was performed with R.
The document was managed with the `bookdown` package for R, built to PDF using \LaTeX.
<!-- A hard copy of the thesis can be found in the University of Wisconsin library system. -->

<!-- The computational environment that was used to generate this version is as
follows: -->

```{r session-info, eval = FALSE, echo = FALSE, include = FALSE}
devtools::session_info()
```


