---
title: "Do Primaries Work? Partisan Representation in Nomination Contests"
subtitle: (Selections from Chapters 2 and 4)
author: "Michael DeCrescenzo"
date: "February 17, 2020"
output:
  xaringan::moon_reader:
    lib_dir: libs
    selfcontained: true
    css: xaringan-themer.css
    # seal: false
    nature:
      ratio: "16:9"
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      beforeInit: "https://platform.twitter.com/widgets.js"
# bibliography: ../../paper-rmd/assets/voter-id-bib.bib
# biblio-style: ../../paper-rmd/assets/apsr-leeper.bst
---

```{r inf, include = FALSE, cache = FALSE, eval = FALSE}
xaringan::inf_mr(here::here("present", "apw-2020", "apw-pos-2020.Rmd"))
# servr::daemon_stop(1)
```


```{r chunks, include = FALSE}
# chunks:
# hide code and messages
# cache everything
knitr::opts_chunk$set(
  eval = TRUE, 
  echo = FALSE, collapse = FALSE,
  include = FALSE, warning = FALSE, message = FALSE,
  cache = TRUE, 
  dpi = 100, fig.retina = 3, 
  fig.align = "center"
)
```



```{r, cache = FALSE}
# Xaringan settings
library("xaringan")
library("xaringanthemer")


# theme colors
primary <- 
  viridis::viridis(1, alpha = 1, begin = 0.35, end = 0.35, direction = 1)
secondary <- 
  viridis::viridis(1, alpha = 1, begin = 0.5, end = 0.5, direction = 1)

primary_light <- xaringanthemer::lighten_color(primary, 0.9)
secondary_light <- xaringanthemer::lighten_color(secondary, 0.5)

yel <- viridis::magma(1, alpha = 1, begin = 0.9, end = 0.9, direction = 1)
purp <-  viridis::magma(1, alpha = 1, begin = 0.5, end = 0.5, direction = 1)

# slide accent colors
black <- xaringanthemer::darken_color(primary, 0.3)
white <- primary_light


mono_light(
  base_color = primary,
  header_font_google = google_font("Source Sans Pro"),
  text_font_google = google_font("Source Serif Pro"), 
  code_font_google = google_font("Inconsolata"), 
  text_bold_color = secondary,
  # code_inline_background_color    = "#F5F5F5", 
  # table_row_even_background_color = "white", 
  extra_css = list(
    "h1, h2, h3" = list("font-weight" = "bold"),
    ".remark-slide-content" = list("font-size" = "24px"),
    ".remark-slide-number" = list("display" = "none"),
    ".remark-inline-code" = list(
      "background" = "#F5F5F5", 
      # "background" = "#e7e8e2", # /* darker */
      "border-radius" = "3px", 
      "padding" = "4px"
    ),
    ".left-code" = list(
      "width" = "38%", "height" = "92%", "float" = "left"
    ),
    ".right-plot" = list(
      "width" = "60%", "float" = "right", "padding-left" = "1%"
    )
  )
)
```

```{r packages, cache = FALSE}
library("here")
library("magrittr")
library("tidyverse")
library("ggdag")
library("patchwork")
filter <- dplyr::filter
```

```{r graphics-theme, cache = FALSE}
# graphics theme
font_fam <- "Source Sans Pro"

# the two parties
dblue <- "#179ee0"
rred <- "#ff5d40"

# for scale_color_*
party_factor_colors <- c("1" = dblue, "2" = rred)
party_code_colors <- c("D" = dblue, "R" = rred)
party_colors <- c("Democrats" = dblue, "Republicans" = rred)


theme_mgd <- function() {
  ggthemes::theme_base(base_size = 14, base_family = font_fam) %+replace% ggplot2::theme(
      legend.background = element_rect(fill = white),
      legend.key = element_rect(fill = white),
      strip.background = element_blank(),
      panel.background = element_rect(fill = white),
      plot.background = element_rect(fill = white, color = NA), 
      axis.ticks = element_line(lineend = "square"), 
      axis.ticks.length = unit(0.15, "lines")
    )
}


ggplot2::theme_set(theme_mgd())

theme_mgd_dag <- function() {
  theme_mgd() %+replace%
  ggplot2::theme(
    panel.border = ggplot2::element_blank(),
    panel.background = ggplot2::element_blank(),
    # axis.line = ggplot2::element_blank(),
    axis.title = ggplot2::element_blank(),
    axis.text = ggplot2::element_blank(),
    axis.ticks = ggplot2::element_blank()
  )
}


update_geom_defaults(
  "text", 
  list(family = font_fam, size = 4, color = black)
)
```


```{r}
img_path <- file.path("present", "apw-2020", "img")
```




class: middle, center

## Note on feedback

???

- defend in summer, industry
- needs to be _defensible_
- I want it to be good, I won't do bad work
- My goal is for it to represent my interests and my record of contributions as a student
- but it doesn't need to have quite the same flavor profile as a project that is aiming for the JOP

???

- Big slides
    - how this started
        - 
    - contours of argument
    - outline of the project
    - non-academic job
        - the purpose of the dissertation is to be a record of my education
        - funky chapter outline
        - I want it to be good, but I don't care about moves that make ideas more "publishable" for its own sake
    - Argument
        - We don't know if theory is true
        - Plausible reasons to doubt theory
        - Existing research doesn't operationalize or uses an inappropriate proxy
    - Inappropriate proxy
    - Modeling concept
        - why do we need this?
        - not the same as primary voters
        - candidates lack exact knowledge of ideological selection into voting
        - What are the groups?
    - Model


---

# What do primaries do?

Why so few moderate politicians (e.g. House members)? 

- Polarization and the "strategic positioning dilemma" (Brady, Han, and Pope 2007)
- Competitive districts $\rightarrow$ moderate legislators (Ansolabehere, Snyder, and Stewart 2001)
- Presence of a primary challenger (Burden 2004)

???

Whither convergence?

- SPD: candidates are scared of primary threat, so they take non-median positions
    - Downs 
- in safer districts, legislators appear "freer" to take more partisan stances
    - Aldrich and other party incentives
- also find that presence of primary challengers are an outward pushing force as well


--

.pull-left[

Studies of primary "mechanisms" aren't clear

- Information costs
- primary turnout, primary "threat"
- open/closed primaries
- groups? interests?

]

???

- If candidates are scared of ideological primary challengers in particular, the threat is that voters can detect when a candidate isn't ideologically pure enough for them. **high information costs**
- primary turnout an indicator of the amount of competitive mobilization happening in primaries, but unrelated to candidate positioning
- parties that are "open" to voters from all parties should encourage moderation? There's basically zero evidence demonstrating
- Is something else going on? Are policy groups at work promoting certain candidates, and the ideological representation of partisans within the district is incidental


--

.pull-right[
Studies of primary mechanisms are weird?

- Presidential vote as ideological proxy, fixed across parties
- **Something huge is missing**
]

???

It seems like a lot of these hypotheses about how primaries affect candidate position-taking assume that voter preferences matter, but voter preferences are not measured in a way that makes sense with these theories

- Presidential vote can't break the parties into separate groups. I don't know if a red district is conservative altogether or if there are just a lot of Republicans
- Composition vs. Disposition problem



---

class: center

## What do votes tell us about ideology?

???

Motivate the problem with current research and how I plan to address item

--

```{r}
mean_ideo <- 
  read_rds(here("present", "apw-2020", "easy-data", "ideo-vote.rds")) %>%
  print()
```

```{r, include = TRUE, fig.width = 8, fig.height = 3.5, out.width = "100%"}
outcome_year <- 2018

ideo_plot_data <- mean_ideo %>%
  filter(is.na(party_group) == FALSE) %>%
  filter(cycle == outcome_year)

ideo_plot <- 
  ggplot(ideo_plot_data) +
  aes(x = district_pres_vs, y = mean_ideo, color = party) +
  facet_wrap(~ str_glue("CCES: {party_group}")) +
  scale_color_manual(
    values = c("All" = "gray", "Democrat" = dblue, "Republican" = rred)
  ) +
  scale_y_continuous(
    breaks = 1:5, 
    labels = c("Very Liberal", "Liberal", "Moderate", 
               "Conservative", "Very Conservative"),
    limits = c(1, 5)
  ) +
  scale_x_continuous(
    limits = c(0, 1),
    labels = scales::percent_format(1)
  ) +
  theme(legend.position = "none") +
  labs(
    x = str_glue("District Republican Vote Share, {outcome_year - 2}"),
    y = str_glue("District Mean Ideological\nSelf-Placement, {outcome_year}")
  )

ideo_plot
```

???

- Vertical axis: ideological self-placement, avg'd within congressional district
- Horizontal axis: the conventionally used proxy. How did the district vote in the prior presidential election



---

class: center

## What do votes tell us about ideology?

```{r, include = TRUE, fig.width = 8, fig.height = 3.5, out.width = "100%"}
ideo_plot + 
  geom_point(
    data = filter(ideo_plot_data, party_group == "All Respondents"),
    shape = 16, alpha = 0.5, size = 2
  ) +
  geom_smooth(
    data = filter(ideo_plot_data, party_group == "All Respondents"),
    method = "lm", aes(group = party), color = "black"
  )
```

???

- Left: expect a positive relationship. But is this because red districts are more conservative? Or just more Republicans? So you break it out by party




---

class: center

## What do votes tell us about ideology?

```{r, include = TRUE, fig.width = 8, fig.height = 3.5, out.width = "100%"}
ideo_plot + 
  geom_point(
    shape = 16, alpha = 0.5, size = 2
  ) +
  geom_smooth(
    method = "lm", aes(group = party), color = "black"
  )
```

???


- Right: and what you see is that, within party, this relationship is much weaker.

What does using the presidential vote as a proxy for ideology within parties: the ideology that's relevant to these hypotheses about primary elections? Very little. 

In red districts, the Republicans aren't much more conservative on average. In blue districts, the Democrats are NO more liberal. (at least in self-categorization)

**So I want to improve the way we measure constituent preferences in the study of primary elections**. The district presidential vote is important as a measure of _general election competition_, but that is only half of the "strategic positioning dilemma." The other half is the competition within the party over the ideological temperature of the nominee. And that's where I'm looking.

---

class: center, inverse

<br><br>

# Dissertation Objectives

--

### Model policy ideology within district-party groups

???

Use survey data on policy questions to model the policy conservatism of Republican partisans in every district, and Democratic partisans in every district, as separate groups.

Not the same as self-placement, as we saw before. Not the same as "symbolic ideology." It's a low-dimensional summary of policy preferences.

--

### Does policy ideology matter for candidate positioning?

???

If I'm a Democratic candidate running in a district where the Democrats are more liberal, am I running on a more liberal platform? And how is that different from running in a district that is safely blue? Those things are different.

--

### Does policy ideology matter for voting in primaries?

???

Do voters pick up on it? In districts where Republicans are more conservative, is the more conservative primary candidate more likely to be nominated


---


```{r spatial-data}
dd <- tibble(
  Left = -0.9,
  Right = 1,
  Respondent = 0.3
) 

long_dd <- dd %>%
  gather(key = label, value = location) %>%
  mutate(height = 0) %>%  
  print()

x_min <- -2
x_max <- 2
u_scale <- 2

utility <- tibble(
  x = seq(x_min, x_max, .01),
  utility_loss = u_scale * -(x - dd$Respondent)^2,
  Right = u_scale * -(x - dd$Right)^2,
  Left = u_scale * -(x - dd$Left)^2
) %>%
  print(n = nrow(.))

```


## Ideal Point Models


Survey respondent prefers policy alternative as $f\left(\text{ideological proximity}\right) + \text{error}$

```{r plot-space-1, out.width = "70%", fig.height = 1.25, fig.width = 5, include = TRUE}
ggplot(long_dd) +
  aes(x = location, y = height) +
  geom_point() +
  annotate(geom = "line", x = c(-2, 2), y = 0) +
  geom_text(aes(label = label), nudge_y = 1) +
  coord_cartesian(ylim = c(-3, 1.25)) +
  geom_line(
    data = utility,
    aes(x = x, y = utility_loss)
  ) +
  theme(
    legend.position = "none",
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.text = element_blank(),
    axis.title = element_blank(),
    axis.ticks = element_blank()
  ) +
  annotate(geom = "linerange",
    x = dd$Right, ymax = 0, ymin = -u_scale*(dd$Respondent - dd$Right)^2,
    linetype = "dotted"
  ) +
  annotate(geom = "linerange",
    x = dd$Left, ymax = 0, ymin = -u_scale*(dd$Respondent - dd$Left)^2,
    linetype = "dotted"
  ) +
  geom_point() +
  NULL
```

--

.pull-left[
$\displaystyle \mathrm{Pr}\left(y_{ij} = \mathrm{Right}\right) = \Phi\left(\frac{\left[ \theta_{i} - \kappa_{j} \right]}{\sigma_{j}} \right)$
]

.pull-right[

Respondent's **ideal point** $\theta_{i}$

**Item midpoint** $\kappa_{j}$

**Item dispersion** $\sigma_{j}$, how "ideological" is this policy choice

]


---

## Group-Level Model

1. Define a **group** $g$ as the intersection of congressional district & party <br>
( $435$ districts $d$ $\times$ $2$ parties $p$ = $870$ groups )

--

2. Assume individual's ideology is **Normal draw**  from group:  $\theta_{i} \sim \mathrm{Normal}\left(\bar{\theta}_{g[i]}, \sigma_{g}\right)$

--

3. Probability that randomly selected $y_{ij} = \mathrm{Right}$: $\quad \displaystyle \bar{\pi}_{gj} = \Phi\left(\frac{\left[ \bar{\theta}_{g} - \kappa_{j} \right]}{ \sqrt{\sigma_{j}^{2} + \sigma_{g}^{2}} } \right)$

--

4. Hierarchical model for group means, smooth estimates across space (and eventually: time)

???


Hierarchical models are a huge amount of work

- Non-centered parameterization to prevent MCMC problems (This is NEW to this literature)


--

5. Hierarchical model for item parameters


???


Item model

- not done in prior group models
- parameterization of this beast is also using cutting edge practices for covariance matrix priors


---

## What you didn't read about the model

???

Model is a lot of work, most of it you aren't seeing in the draft

--

.left-code[

- Weighted outcome data $\rightarrow$ custom quasilikelihood function

- Parameterizations of hierarchical models for Hamiltonian Monte Carlo

- Detailed look at priors (& contrast with previous work)

]




--

.right-plot[
```{r}
prior_probit <- 
  tibble(
    x = seq(-5, 5, .01),
    Probability = pnorm(x)
  ) %>%
  print()
```

```{r, include = TRUE, out.width = "100%", fig.width = 4.5, fig.height = 3.5}
ggplot(prior_probit) +
  aes(x = x, y = Probability) +
  labs(
    title = "Managing Prior Expectations",
    subtitle = "for a Probit Model",
    x = "Latent Index Value", y = "Success Probability"
  ) +
  coord_cartesian(ylim = c(0, 1)) +
  scale_x_continuous(breaks = seq(-4, 4, 2)) +
  # layered ribbons
  geom_ribbon(
    data = tibble(
      x = c(qnorm(.01), qnorm(.99)),
      ymin = -2, ymax = 2
    ),
    aes(ymin = ymin, ymax = ymax, y = NULL),
    fill = secondary_light, alpha = 0.3
  ) +
  geom_ribbon(
    data = tibble(
      x = c(qnorm(.025), qnorm(.975)),
      ymin = -2, ymax = 2
    ),
    aes(ymin = ymin, ymax = ymax, y = NULL),
    fill = secondary_light, alpha = 0.5
  ) +
  geom_ribbon(
    data = tibble(
      x = c(qnorm(.05), qnorm(.95)),
      ymin = -2, ymax = 2
    ),
    aes(ymin = ymin, ymax = ymax, y = NULL),
    fill = secondary_light, alpha = 0.7
  ) +
  # 90%
  annotate(geom = "text", x = 3, y = 0.5, hjust = -0.1, label = "Inner 90%") +
  annotate(
    geom = "segment", 
    x = 3, xend = 0.85, y = 0.50, yend = 0.50
  ) +
  annotate(geom = "point", x = 0.85, y = 0.5, size = 1) +
  # 95%
  annotate(geom = "text", x = 3, y = 0.35, hjust = -0.1, label = "Inner 95%") +
  annotate(
    geom = "segment", 
    x = 3, xend = mean(qnorm(c(.975, .95))), y = 0.35, yend = 0.35
  ) +
  annotate(geom = "point", x = mean(qnorm(c(.975, .95))), y = 0.35, size = 1) +
  # 99%
  annotate(geom = "text", x = 3, y = .20, hjust = -0.1, label = "Inner 98%") +
  annotate(
    geom = "segment", 
    x = 3, xend = mean(qnorm(c(.99, .975))), y = .20, yend = .20
  ) +
  annotate(geom = "point", x = mean(qnorm(c(.99, .975))), y = 0.2, size = 1) +
  # cdf
  geom_line(size = 0.75) +
  annotate(
    geom = "text", label = "Normal CDF",
    x = -2.5, y = .85,
    hjust = 1.1
  ) +
  annotate(
    geom = "segment", x = -2.5, xend = qnorm(.85),
    y = .85, yend = .85
  ) +
  annotate(geom = "point", x = qnorm(.85), y = .85, size = 1)
```
]

---

class: middle

```{r read-thetas}
thetas <- read_rds(here("present", "apw-2020", "easy-data", "theta-intervals.rds"))
```


.left-code[
## Group Ideal Points

- ANES & CCES data, 2012–2018

- Location and scale 

- Tbd: post-process ideal points

- Homoskedasticity assumptions

- Very little correlation across parties

]



.right-plot[

```{r theta-intervals, include = TRUE, out.width = "100%", fig.width = 5, fig.height = 4}
ggplot(thetas) +
  aes(x = party_rank,  y = estimate, color = as.factor(party)) +
  geom_linerange(
    aes(ymin = conf.low_0.95, ymax = conf.high_0.95),
    show.legend = FALSE, size = 0.25, alpha = 0.25
  ) +
  geom_linerange(
    aes(ymin = conf.low_0.9, ymax = conf.high_0.9),
    show.legend = FALSE, alpha = 0.35
  ) +
  geom_linerange(
    aes(ymin = conf.low_0.5, ymax = conf.high_0.5),
    show.legend = FALSE
  ) +
  geom_point(size = 0.5, color = "black") +
  scale_color_manual(values = party_factor_colors) +
  # coord_flip() +
  # scale_x_reverse() +
  labs(
    x = "Distric-Party Groups\nOrdered Most Liberal to Most Conservative", 
    y = "Group Ideal Point",
    title = "District-Party Ideal Point Estimates",
    subtitle = "Two Parties x 435 Districts"
  ) + 
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()
  ) +
  annotate("text", 
    y = c(-0.7, 0.7), x = c(100, 100),
    label = c("Democrats", "Republicans")
  )
```

]


---

class: center, middle, inverse

# How does group ideology affect candidate positioning?

???

This is a literature that is largely regression focused, a lot of selection on observables as key identifying assumptions.

Given the way the estimates are constructed, I haven't yet come up with a genius way to break out of selection on observables either. In fact the way the estimates are constructed, estimates are all borrowing information from one another, so we're limited in the kind of natural experiment/RDD/instrumental variables things that you might think of.

So I'm trying to do as best as I can given these constraints. I don't want to just plug things into a regression and look at a coefficient. I want to think hard about what is the estimand that we're interested in, what does the causal model/causal graph allowing me to make inferences about?

So that's why I take the particular causal inference/causal graphs approach that I take even if we can't get out of selection on observables.


---

## How does group ideology affect candidate positioning?

.left-code[

**Strategic positioning dilemma**: "primary constituency" or "district competition"?


Effect of partisan preferences, _controlling for_ district competition

Confounding problem: party median affects district median
]


.right-plot[

```{r problem-dag, include = TRUE, fig.height = 3, fig.width = 5, out.width = "100%"}
problem_dag <- 
  dagify(
    C ~ T + V + U,
    V ~ T + U,
    exposure = "T",
    outcome = "C",
    coords = tribble(
      ~ name,      ~ x,    ~ y,
      "T",      0,      0,
      "C",      2,      0,
      "V",      1,      1,
      "U",      2,      1
    ),
    labels = c(
      "T" = "Partisan Ideology",
      "C" = "Candidate Positioning",
      "V" = "District Voting"
    )
  ) %>%
  tidy_dagitty() %>%
  # node_parents("C") %>%
  mutate(
    pt_label = case_when(
      name == "T" ~ "Group Ideology",
      name == "C" ~ "Candidate DIME",
      name == "V" ~ "District Vote",
      name == "U" ~ "Confounders"
    )
  )

ggplot(problem_dag) +
  aes(x = x, y = y, xend = xend, yend = yend) +
  geom_dag_edges() +
  geom_dag_point(
    aes(color = (name == "U")), show.legend = FALSE
  ) +
  geom_dag_label_repel(
    aes(label = pt_label), 
    family = font_fam,
    segment.color = "black"
  ) +
  scale_color_manual(values = c("TRUE" = secondary, "FALSE" = "black")) +
  theme_mgd_dag() +
  expand_plot(
    expand_x = expand_scale(c(0.2, 0.2)), 
    expand_y = expand_scale(c(0.2, 0.2))
  ) +
  NULL  
```

]

???

**What are the data**


---

### Average Controlled Direct Effects using Sequential-_g_

Special parametric case of **structural nested mean model**

```{r create-g-dag}
g_dag <- 
  dagify(
    C ~ T + V + Z + X,
    V ~ T + X + Z,
    Z ~ T + X,
    T ~ X,
    exposure = "T", 
    outcome = "C",
    coords = tribble(
      ~ name,  ~ x, ~ y,
      "T",       1,   2,
      "C",       3,   1,
      "Z",       1,   0,
      "V",       2,   0,
      "X",       0,   1
    )
  ) %>%
  tidy_dagitty(layout = "auto") %>%
  mutate(
    label = case_when(
      name == "V" ~ "Past Presidential Vote",
      name == "T" ~ "Partisan Ideology",
      name == "C" ~ "Candidate Position",
      name == "X" ~ "Pre-Treatment Confounders",
      name == "Z" ~ "Intermediate Confounders"
    )
  ) %>%
  as_tibble() %>%
  print()
```



```{r}
base_dag <- g_dag %>%
  ggplot() +
  aes(x = x, y = y, xend = xend, yend = yend) +
  coord_cartesian(
    ylim = c(-0.25, 2.25),
    xlim = c(-0.25, 3.25)
  ) +
  # theme_dag() +
  theme_mgd_dag() +
  NULL

mediator_dag <- base_dag +
  geom_dag_point() +
  geom_dag_edges() +
  geom_dag_text(aes(label = name)) +
  labs(title = "\nStage 1", subtitle = "Mediator-outcome model\n") +
  NULL

exposure_dag <- base_dag +
  geom_dag_point(aes(color = name %in% c("Z", "V"))) +
  geom_dag_edges(
    data_directed = g_dag %>%
    filter(
      (name %in% c("V", "Z") & to == "C") == FALSE,
      to %in% c("V", "Z") == FALSE
    )
  ) +
  geom_dag_edges(
    data_directed = g_dag %>%
      filter(
        to %in% c("Z", "V") | (name == "Z" & to == "C"), 
        name %in% c("U1", "U2") == FALSE
      ),
    edge_linetype = "dashed", 
    edge_color = "gray50"
  ) +
  geom_dag_text(aes(label = ifelse(name == "C", "b(C)", name))) +
  scale_color_manual(values = c("TRUE" = primary, "FALSE" = "black")) +
  labs(
    title = "\nStage 2", 
    subtitle = "Exposure-outcome model\nwith demediated outcome"
  ) +
  theme(legend.position = "none")
```


```{r include = TRUE, out.width = "100%", fig.height = 5, fig.width = 11}
mediator_dag + exposure_dag 
```


???

What are the nodes

What gets identified

Time sequencing

- one set of ideal points, multiple years of data
- some slippage: mediation happens in reverse time


---

## Turn up, tune in, blip down

--

Obtain **demediation/blipdown function** from first stage

--

Blipdown function: 
$\psi_{i}(\delta, v_{g[i]}, \tilde{v}) = \delta \times \left(\begin{bmatrix} v_{g[1]} \\ v_{g[2]} \\ \vdots \\ v_{g[n]} \end{bmatrix} - \tilde{v} \right)$

--


Demediate outcome: $\begin{bmatrix} y_1 \\ y_2 \\ \vdots \\ y_n \end{bmatrix} - \boldsymbol{\mathbf{\psi}}(\delta, v_{g[i]}, \tilde{v}) = \begin{bmatrix} b(y_1)  \\ b(y_2) \\ \vdots \\ b(y_n) \end{bmatrix}$



---

# Propagating Uncertainty

Sources of uncertainty: ideal point model, stage 1 model (demediation), stage 2 model (ACDE)

--

Estimation routine:

- Sample **one** MCMC iteration from ideal point model.

--

- Estimate stage 1 model, district vote (mediator) effect. 

--

- Sample **one** coefficient draw of mediator effect. Demediate candidate ideal points.

--

- Estimate stage 2 model, group ideology (treatment) effect. Sample **one** ACDE draw.

--

- Repeat 1000 times and combine samples.


---

## Key findings: ACDE

```{r read-g}
g_samples <- 
  here("data", "model-output", "04-positioning", "pos-g-samples.rds") %>%
  read_rds() %>%
  print()
```

```{r unnest-samples}
mediator_samples <- g_samples %>%
  select(party, incumbency, .draw, mediator_samples) %>%
  unnest(cols = mediator_samples) %>%
  ungroup() %>%
  mutate(
    incumbency = str_glue("{incumbency}s")
  ) %>%
  print()

direct_samples <- g_samples %>%
  select(party, incumbency, .draw, direct_samples) %>%
  unnest(cols = direct_samples) %>%
  ungroup() %>%
  mutate(
    incumbency = str_glue("{incumbency}s")
  ) %>%
  print()

total_samples <- g_samples %>%
  select(party, incumbency, .draw, direct_samples, total_samples) %>%
  unnest(cols = c(total_samples, direct_samples)) %>%
  ungroup() %>%
  mutate(
    indirect_effect = total_effect - direct_effect,
    incumbency = str_glue("{incumbency}s")
  ) %>%
  print()
```

```{r g-summaries}
mediator_summary <- mediator_samples %>%
  group_by(party, incumbency) %>% 
  summarize(
    sample_mean = mean(mediator_effect), 
    conf.low = quantile(mediator_effect, .05), 
    conf.high = quantile(mediator_effect, .95),
    n_samples = n()
  ) %>%
  print()

direct_summary <- direct_samples %>%
  group_by(party, incumbency) %>% 
  summarize(
    sample_mean = mean(direct_effect), 
    conf.low = quantile(direct_effect, .05), 
    conf.high = quantile(direct_effect, .95),
    n_samples = n()
  ) %>%
  print()

total_summary <- total_samples %>%
  group_by(party, incumbency) %>% 
  summarize(
    sample_mean = mean(total_effect), 
    conf.low = quantile(total_effect, .05), 
    conf.high = quantile(total_effect, .95),
    n_samples = n()
  ) %>%
  print()

indirect_summary <- total_samples %>%
  group_by(party, incumbency) %>% 
  summarize(
    sample_mean = mean(total_effect - direct_effect), 
    conf.low = quantile(total_effect - direct_effect, .05), 
    conf.high = quantile(total_effect - direct_effect, .95),
    n_samples = n()
  ) %>%
  print()
```


```{r plot-direct, include = TRUE, fig.width = 8, fig.height = 3, out.width = "100%"}
ggplot(direct_samples) +
  aes(
    x = direct_effect, y = as.factor(party), 
    fill = as.factor(party), color = as.factor(party)
  ) +
  geom_vline(
    xintercept = 0, color = "gray", size = 0.5, linetype = "dashed"
  ) +
  # annotate(
  #   geom = "segment", x = 0, xend = 0, y = 0, yend = 3.25,
  #   color = "gray"
  # ) +
  geom_segment(
    data = direct_summary,
    aes(x = conf.low, xend = conf.high, 
        y = as.factor(party), yend = as.factor(party))
  ) +
  geom_point(
    data = direct_summary, aes(x = sample_mean, y = as.factor(party))
  ) +
  ggridges::geom_ridgeline(
    stat = "binline", draw_baseline = FALSE,
    boundary = 0, bins = 40, scale = 0.4,
    alpha = 0.25
  ) +
  facet_wrap(~ incumbency, nrow = 1, strip.position = "bottom") +
  scale_color_manual(values = party_factor_colors) +
  scale_fill_manual(values = party_factor_colors) +
  scale_y_discrete(
    breaks = c(1, 2), labels = c("Democrats", "Republicans")
  ) +
  theme(
    legend.position = "none",
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.line.x = element_line(),
    axis.ticks.y = element_blank(),
    strip.placement = "outside",
    # axis.text.y = element_blank()
  ) +
  coord_cartesian(ylim = c(1.35, 2.25)) +
  labs(
    title = "Effects of District-Party Ideology",
    subtitle = "Controlled Direct Effect on Candidate Dime Score",
    y = NULL,
    x = NULL
  ) +
  NULL
```

--

- Within party sd(Dime) $\approx 0.5$, within party sd( $\bar{\theta}_{g}$ ) $\approx 0.2$
--

- One sd change in district-party ideal points $\approx$ $0.2$ sd effect on Dime score

???

Dime sd within party is about 0.5,

theta sd within party is about 0.2

A ONE POINT increase in theta is ~ 1 sd effect on DIME

A one-sd increase in theta is ~ a fifth sd increase on DIME, which feels big

---

## What's next?

.pull-left[
#### Model Chapter

- Dynamic model

- More data, farther back in time

- Covariate adjustment

]

.pull-right[
#### Candidate Positioning Chapter

- Primary institutions (caucuses, conventions, openness) and effect modification

- Covariate adjustment

- Flatten the hierarchy

]





---

class: center, middle, inverse

# Do Primaries Work?

## Partisan Representation in Nomination Contests